---
title: "Linear Regression"
author: "Krishna Nadiminti"
date: "January 19, 2016"
output: html_document
---


##Reading and Cleaning the Data

```{r}
docML1 <- read.csv("docML1.csv")
docML1<- subset(docML1,!(is.na(docML1$dwnlds)))
hr<-docML1$hour
dwnl<-docML1$dwnlds
### Replacing Nan with 0
#dwnl<-replace(dwnl, is.na(dwnl), 0)
```


##Plotting the Elements as Scatter Plot

```{r}

plot(dwnl~hr)

```


##Computing Least Square Regression Equation

$$Y=mX + C$$

$$m= \sum\limits_{i=1}^{N}({x}_i-\bar{x})({y}_i-\bar{y}_i)/\sum\limits_{i=1}^{N}({x}_i-\bar{x})^2$$

```{r}

Sxx=0;
Sxy=0;
for(i in 1:length(hr)) {
    
   Sxx=Sxx+((hr[i]-mean(hr))*(dwnl[i]-mean(dwnl)))
   Sxy= Sxy+((hr[i]-mean(hr))^2)
    
}
 
```


####Slope for the Linear regression is


```{r}
(slope<- (Sxx/Sxy))

```


#### Intercept for the Linear Regression Equation is

```{r}
(intercept<- mean(dwnl)-(slope*mean(hr)))
```

#### Predicted Downloads on noon Fifth day of next month

No of Hours on of Fifth day is 852

```{r}
(Respone<- slope*852+intercept)

```

Total number of Downloads at 852 hours is 3236.685

#### Visualizing the Regression Equation on the scatterplot

```{r}
plot(dwnl~hr)
abline(intercept,slope)
```

Upon Plotting the Scatterplot we can consider that the popularity of the book has been steadily increasing over the period of time and the variables are nearly linear fashioned.

##Comparing the results with Existing LinearModel

```{r}
modelLinear<- lm(dwnlds~hour,data = docML1)
summary(modelLinear)
```

Seems both the Functions have yielded the similar Results